---
title: "MA_thesis_analysis"
output: html_document
author: Line Kruse 
---

```{r setup, include=FALSE}
library(pacman)
p_load(psychonetrics, dplyr, qgraph, ggplot2, bootnet, Rmpfr, semPlot, GGally, corpcor, lvnet, BGGM, pracma, semPlot, matrixcalc, mctest, factoextra, readxl)
```

#Import data 

```{r}
#SIPS symptoms 
clinical <- read.delim("/path/PNC_clinical_data_cleaned.txt", sep=" ",header=TRUE)

#Cognitive tasks
cognitive <- read.delim("/path/PNC_cognitive_data_cleaned.txt", sep=" ",header=TRUE)

#Genetic cluster weights derived from NBS analysis 
cluster_weights <- read.delim("/path/cluster_weights.txt", sep=" ",header=TRUE)

#Demographics
demo <- read.delim("/path/clin_info_n6481.txt", sep=",",header=TRUE)

demo2 <- read_excel("/path/fullSample_behavioral-tapfixDA.xlsx")

```

#Clean and combine datasets 

```{r}
#Rename ID column in cluster df 
cluster_weights <- rename(cluster_weights, SUBJID="ID")

#Merge dataframes 
data <- merge(clinical,cognitive, by="SUBJID")
data <- merge(data, cluster_weights, by="SUBJID")
```

#Demographics and basic statistics 

```{r}
#Get demographics stats - Age and Sex
demo <- demo[,c(2:4)]
demo_df <- merge(data, demo, by="SUBJID")
summary(demo_df$age_at_cnb)
sd(demo_df$age_at_cnb)
min(demo_df$age_at_cnb)  
max(demo_df$age_at_cnb) 
table(demo_df$Sex) 

#Get demographics - Race 
demo2 <- demo2[,c(1:3)]
IDs <- data[,c(1,2)]
demo2 <- merge(demo2, IDs, by="SUBJID")
summary(demo2$Race)

#Check number of subjects in each cluster (based on highest value)
cluster_subset <- data_remNA[,c(20:22)]
x <- colnames(cluster_subset)[apply(cluster_subset, 1, which.max)]
summary(x)

#Calculate mean and sd for "weights" in each cluster (do before sclaing)
colMeans(cluster_subset) 
apply(cluster_subset, 2, sd)

#Mean symptom scores  
SIP_sum <- rowSums(data[,c(2:8)])
data$SIP_sum <- SIP_sum
summary(data$SIP_sum)
sd(data$SIP_sum, na.rm = T)

#Mean WRAT scores 
summary(data$WRAT)
sd(data$WRAT)

#Split male/female
female <- demo_df[demo_df$Sex=="F",]
male <- demo_df[demo_df$Sex=="M",]

summary(female$age_at_cnb)
sd(female$age_at_cnb)

summary(male$age_at_cnb)
sd(male$age_at_cnb)

summary(female$SIP_sum)
sd(female$SIP_sum, na.rm = T)

summary(male$SIP_sum)
sd(male$SIP_sum, na.rm=T)

summary(female$WRAT)
sd(female$WRAT)

summary(male$WRAT)
sd(male$WRAT)
```

#Rename variables, log-transform and scale 

```{r}
#Remove ID column 
data <- data[,-1]

#Rename vars 
data <- data %>%
  rename(Attentional_Deficits="SIP001",
         Disorganized_Speech="SIP030",
         PerceptionSelfOthersWorld="SIP035",
         Emotional_Expression="SIP037",
         Occupational_Functioning="SIP041",
         Avolition="SIP043",
         Psychosis="PSY",
         Social_Cognition="PADT",
         Facial_Memory="FMT",
         Emotion_Recognition="PEIT",
         Verbal_Memory="PWMT",
         Verbal_Reasoning="PVRT",
         Motor_Planning="PMT",
         Abstraction="PMRT",
         Motor_Speed="FT1",
         Visuospatial="VOLT",
         Working_Memory="NBACK",
         Attention="PCPTCR",
         Academic_Skills="WRAT")

#Remove NAs 
data_remNA <- na.omit(data)

#Log transform right skewed data 
data_remNA$sub1 <- log(data_remNA$sub1)
data_remNA$sub2 <- log(data_remNA$sub2)
data_remNA$sub3 <- log(data_remNA$sub3)

#Scale 
data_SIP <- scale(data_remNA)
```

#Principal Component Analysis (PCA)

```{r}
#Perform pca using spectral decomposition 
pca_sd <- princomp(data_SIP, cor=T, scores=T)
summary(pca_sd) 

#Inspect components
res_var <- get_pca_var(pca_sd)
res_var$contrib 
res_var$cos2 

#Inspect eigenvalues
eig <- get_eigenvalue(pca_sd)

#Inspect variance
summary(pca_sd)

#Plot principal components (scree plot)
pdf(file="/path/PCA_screePlot.pdf", width=5, height=3)
fviz_eig(pca_sd)
dev.off()

plot(pca_sd)
```

#Exploratory Factor Analysis (EFA)
##With 2, 3, 4 and 5 factors 

```{r}
#Model estimation 
efa_2f <- factanal(data_SIP, factors=2, rotation="promax",scores="regression")
efa_2f_loadings <- loadings(efa_2f)

efa_3f <- factanal(data_SIP, factors=3, rotation="promax",scores="regression")
efa_3f_loadings <- loadings(efa_3f)

efa_4f <- factanal(data_SIP, factors=4, rotation="promax",scores="regression")
efa_4f_loadings <- loadings(efa_4f)

efa_5f <- factanal(data_SIP, factors=5, rotation="promax",scores="regression")
efa_5f_loadings <- loadings(efa_5f)

#Inspect factor loadings 
pdf(file="/path/EFA_2f_loadings.pdf")
qgraph.loadings(efa_2f_loadings, rotation="promax",minimum=0.2,cut=0.4,vsize=c(5,15),borders=T, labels=c(1:22), layout="tree2", title="EFA: Two-factor model loadings")
dev.off()

pdf(file="/path/EFA_3f_loadings.pdf")
qgraph.loadings(efa_3f_loadings, rotation="promax",minimum=0.2,cut=0.4,vsize=c(5,15),borders=T,labels=c(1:22), layout="tree2", title="EFA: Three-factor model loadings")
dev.off()

pdf(file="/path/EFA_4f_loadings.pdf")
qgraph.loadings(efa_4f_loadings, rotation="promax",minimum=0.2,cut=0.4,vsize=c(5,15),borders=T,labels=c(1:20), layout="tree2", title="EFA: Four-factor model loadings")
dev.off()

pdf(file="/path/EFA_5f_loadings.pdf")
qgraph.loadings(efa_5f_loadings, rotation="promax",minimum=0.2,cut=0.4,vsize=c(5,15),borders=T,labels=c(1:20), layout="tree2", title="EFA: Five-factor model loadings")
dev.off()

```

#Create factor loading matrices based on EFA results

```{r}
lambda_2f <- matrix(0,22,2)
lambda_2f[c(3,18,20,21),1] <- 1
lambda_2f[c(14,18,20:22),2] <- 1

lambda_3f <- matrix(0,22,3)
lambda_3f[c(1,8:19),1] <- 1
lambda_3f[c(3,20,21),2] <- 1
lambda_3f[c(20:22),3] <- 1

lambda_4f <- matrix(0,22,4)
lambda_4f[c(1,8:19),1] <- 1
lambda_4f[c(1:7),2] <- 1
lambda_4f[c(20,21),3] <- 1
lambda_4f[c(20:22),4] <- 1

lambda_5f <- matrix(0,22,5)
lambda_5f[c(8:10,12:19),1] <- 1
lambda_5f[c(1:7),2] <- 1
lambda_5f[c(20,21),3] <- 1
lambda_5f[c(8,9,11:14,16:19),4] <- 1
lambda_5f[c(20:22),5] <- 1
```

#Confirmatory Factor Model (CFA)
##Compare fit of the five factor structures derived from EFA 

```{r}
obsvars <- colnames(data_SIP)

#cfa_1f <- lvm(data_SIP, lambda=lambda_1f, vars=obsvars, identification = "variance", latents=1)%>%runmodel

cfa_2f <- lvm(data_SIP, lambda=lambda_2f, vars=obsvars, identification = "variance", latents=c(1,2))%>%runmodel

cfa_3f <- lvm(data_SIP, lambda=lambda_3f, vars=obsvars, identification = "variance", latents=c(1:3))%>%runmodel

cfa_4f <- lvm(data_SIP, lambda=lambda_4f, vars=obsvars, identification = "variance", latents=c(1:4))%>%runmodel

cfa_5f <- lvm(data_SIP, lambda=lambda_5f, vars=obsvars, identification = "variance", latents=c(1:5))%>%runmodel

#Compare fit of the five models - see which one to use in comparison with GGM, LNM and RNM
compare(cfa_2f, cfa_3f, cfa_4f, cfa_5f)

#Posterior model probabilities 
comparison_CFA <- compare(Two_factor=cfa_2f, Three_factor=cfa_3f, Four_factor = cfa_4f, Five_factor=cfa_5f)

BICs_cfa <- mpfr(comparison_CFA$BIC, 100)
BICtrans_cfa <- exp(-0.5 * BICs_cfa) 
sumBICtrans_cfa <- sum(BICtrans_cfa) 
comparison_CFA$modelProbability <- as.numeric(BICtrans_cfa / sumBICtrans_cfa)

pdf(file="/path/CFA_model_comparison_posterior_probability_plot.pdf")
ggplot(comparison_CFA, aes(x=model, y=modelProbability))+
  geom_bar(stat="identity")+
  xlab("")+
  ylab("Posterior model probability, one-factor")+
  theme_bw()+
  labs(title="CFA posterior model probabilities")+
  ylim(0,1)
dev.off()

```

#Define three-factor loading structure (theory based)

```{r}
#Define factor structure 
lambda_3f_theory <- matrix(0,22,3)
lambda_3f_theory[c(1:7),1] <- 1 #Clinical 
lambda_3f_theory[c(8:19),2] <- 1 #Cognitive 
lambda_3f_theory[c(20:22),3] <- 1 #Genetic 

```

#Split data into training and test set 

```{r}
train_data <- data_SIP[c(TRUE, FALSE),] #Odd cases
test_data <- data_SIP[c(FALSE, TRUE),] #Even cases 
```

#Out-of-sample model fit comparison of three- and five-factor CFA 

```{r}
#Five factor CFA 
CFA_train_5f <- lvm(train_data, lambda=lambda_5f, vars=obsvars, identification = "variance", latents=c(1:5))%>%runmodel

CFA_lambda_5f <- getmatrix(CFA_train_5f, "lambda")
CFA_struc_5f <- 1*(CFA_lambda_5f != 0)

CFA_test_5f <- lvm(test_data, lambda=CFA_struc_5f, vars=obsvars, identification = "variance", latents=c(1:5))%>%runmodel

#Three factor CFA 
CFA_train_3f <- lvm(train_data, lambda=lambda_3f_theory, vars=obsvars, identification = "variance", latents=c(1:3))%>%runmodel

CFA_lambda_3f <- getmatrix(CFA_train_3f, "lambda")
CFA_struc_3f <- 1*(CFA_lambda_3f != 0)

CFA_test_3f <- lvm(test_data, lambda=CFA_struc_3f, vars=obsvars, identification = "variance", latents=c(1:3))%>%runmodel

#Comparison 
compare(CFA_test_5f, CFA_test_3f)
```

#Out-of-sample model fit comparison of three- and five-factor LNM 

```{r}
#Five-factor LNM 
LNM_train_5f <- lnm(train_data, lambda=lambda_5f, vars=obsvars, identification="variance", latents=c(1:5))%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

LNM_lambda_5f <- getmatrix(LNM_train_5f, "lambda")
LNM_sigmaZeta_5f <- getmatrix(LNM_train_5f, "sigma_zeta")
LNM_sigmaEpsilon_5f <- getmatrix(LNM_train_5f, "sigma_epsilon")

struc_lambda_5f <- 1*(LNM_lambda_5f != 0)
struc_sigmaZeta_5f <- 1*(LNM_sigmaZeta_5f != 0)
struc_sigmaEpsilon_5f <- 1*(LNM_sigmaEpsilon_5f != 0)

LNM_test_5f <- lnm(test_data, vars=obsvars, identification="variance", lambda=struc_lambda_5f, sigma_zeta=struc_sigmaZeta_5f, sigma_epsilon=struc_sigmaEpsilon_5f, latents=c(1:5))%>%
  runmodel

#Three-factor LNM 
LNM_train_3f <- lnm(train_data, lambda=lambda_3f_theory, vars=obsvars, identification="variance", latents=c(1:3))%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

LNM_lambda_3f <- getmatrix(LNM_train_3f, "lambda")
LNM_sigmaZeta_3f <- getmatrix(LNM_train_3f, "sigma_zeta")
LNM_sigmaEpsilon_3f <- getmatrix(LNM_train_3f, "sigma_epsilon")

LNM_model <- lnm(data_SIP, lambda=lambda_3f_theory, vars=obsvars, identification="variance", latents=c(1:3))%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

LNM_lambda_3f <- getmatrix(LNM_model, "lambda")
LNM_sigmaZeta_3f <- getmatrix(LNM_model, "sigma_zeta")
LNM_sigmaEpsilon_3f <- getmatrix(LNM_model, "sigma_epsilon")

struc_lambda_3f <- 1*(LNM_lambda_3f != 0)
struc_sigmaZeta_3f <- 1*(LNM_sigmaZeta_3f != 0)
struc_sigmaEpsilon_3f <- 1*(LNM_sigmaEpsilon_3f != 0)

LNM_test_3f <- lnm(test_data, vars=obsvars, identification="variance", lambda=struc_lambda_3f, sigma_zeta=struc_sigmaZeta_3f, sigma_epsilon=struc_sigmaEpsilon_3f, latents=c(1:3))%>%
  runmodel

#Comparison 
compare(LNM_test_5f, LNM_test_3f)
```

#Out-of-sample model fit comparison of three- and five-factor RNM 

```{r}
#Five-factor RNM 
RNM_train_5f <- rnm(train_data, lambda=lambda_5f, vars=obsvars, estimator="FIML")%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

RNM_omegaEpsilon_5f <- getmatrix(RNM_train_5f, "omega_epsilon")
RNM_omegaEpsilon_5f[is.na(RNM_omegaEpsilon_5f)] <- 0
RNM_lambda_5f <- getmatrix(RNM_train_5f, "lambda")

RNM_struc_omegaEpsilon_5f <- 1*(RNM_omegaEpsilon_5f != 0)
RNM_struc_lambda_5f <- 1*(RNM_lambda_5f != 0)

RNM_test_5f <- rnm(test_data, lambda=RNM_struc_lambda_5f, omega_epsilon=RNM_struc_omegaEpsilon_5f, vars=obsvars, estimator="FIML")%>%
  runmodel
##### Test model not converging #####

#Three-factor RNM 
RNM_train_3f <- rnm(train_data, lambda=lambda_3f_theory, vars=obsvars, estimator="FIML")%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

RNM_train_3f <- rnm(data_SIP, lambda=lambda_3f_theory, vars=obsvars, estimator="FIML")%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion="bic")

RNM_omegaEpsilon_3f <- getmatrix(RNM_train_3f, "omega_epsilon")
RNM_omegaEpsilon_3f[is.na(RNM_omegaEpsilon_3f)] <- 0
RNM_lambda_3f <- getmatrix(RNM_train_3f, "lambda")

RNM_struc_omegaEpsilon_3f <- 1*(RNM_omegaEpsilon_3f != 0)
RNM_struc_lambda_3f <- 1*(RNM_lambda_3f != 0)

RNM_test_3f <- rnm(test_data, lambda=RNM_struc_lambda_3f, omega_epsilon=RNM_struc_omegaEpsilon_3f, vars=obsvars, estimator="FIML")%>%
  runmodel

RNM_test_3f %>% fit
```

#Gaussian Graphical Model (GGM)
##Out-of-sample fit 

```{r}
GGM_train <- ggm(train_data, vars=obsvars, estimator="FIML")%>%
  runmodel%>%
  prune(adjust="fdr", alpha=0.01)%>%
  modelsearch(verbose=F, criterion = "bic")

GGM_omega <- getmatrix(GGM_train, "omega")
GGM_struc <- 1*(GGM_omega != 0)

GGM_test <- ggm(test_data, vars=obsvars, estimator="FIML", omega=GGM_struc)%>%
  runmodel

GGM_test %>% fit

```

#Out-of-sample model fit comparison of best CFA, LNM, RNM and GGM 

```{r}
comparison_test <- compare(
  CFA=CFA_test_3f, 
  GGM=GGM_test, 
  LNM=LNM_test_3f,
  RNM=RNM_test_3f)

BICs <- mpfr(comparison_test$BIC, 100)
BICtrans <- exp(-0.5 * BICs)
sumBICtrans <- sum(BICtrans)
comparison_test$modelProbability <- as.numeric(BICtrans / sumBICtrans)

pdf(file="/path/All_models_comparison_posterior_probability.pdf")
ggplot(comparison_test, aes(x=model, y=modelProbability))+
  geom_bar(stat="identity")+
  xlab("")+
  ylab("Posterior model probability, one-factor")+
  theme_bw()+
  ylim(0,1)+
  labs(title="2-factor models - Out-of-sample comparison", subtitle = "Posterior Model Probability")
dev.off()
```

#Graphs 

```{r}
p_load(colorspace)

groups <- list("Clinical"=c(1:7),"Cognitive"=c(8:19),"Genetic"=c(20:22))
colors <- hcl.colors(3, palette="Set 2")

#GGM 
pdf(file="/path/GGM.pdf", width=6, height=3)
GGM_net <- getmatrix(GGM_model, "omega")
GGM_graph <- qgraph(GGM_net, theme="colorblind",layout="spring", nodeNames=obsvars,groups=groups,color=colors, legend.cex=0.2)
dev.off()

#CFA 
pdf(file="/path/CFA.pdf", width=5, height=3)
CFA_factorLoadings <- getmatrix(CFA_model_3f_theory, "lambda")
CFA_graph <- qgraph.loadings(CFA_factorLoadings, theme="colorblind", model="reflective",rotation="promax", vsize=c(5,10), asize=3, layout="tree", groups=groups, nodeNames=obsvars, legend.cex=0.4, color=colors)
dev.off()

#LNM 
mod_LNM <- lisrelModel(LY=getmatrix(LNM_model_3f, "lambda"),
                          PS=getmatrix(LNM_model_3f, "sigma_zeta"),
                          TE=getmatrix(LNM_model_3f, "sigma_epsilon"),
                          latNamesEndo = c(1,2,3),
                          manNamesEndo = obsvars)

omega_LNM <- getmatrix(LNM_model_3f, "omega_zeta")

pdf(file="/path/LNM.pdf", width=5, height=3)
LNM_graph <- semPaths(mod_LNM,
                  what="std",
                  whatLabels="omit", #to make labels represent parameter estimates 
                  style="lisrel",
                  residScale=10,
                  theme="colorblind",
                  layout="tree2",
                  #layout="spring",
                  cardinal="lat cov",
                  curvePivot=TRUE,
                  sizeMan=4,
                  sizeLat=8,
                  sizeInt=8,
                  edge.label.cex = 1,
                  mar=c(6,1,8,1),
                  reorder=FALSE,
                  width=8,
                  height=5,
                  groups="latents",
                  pastel=TRUE,
                  borders=FALSE,
                  nCharNodes=0,
                  )
dev.off()

#RNM 
net_RNM <- getmatrix(RNM_model_3f, "omega_epsilon")
facLoadings_RNM <- getmatrix(RNM_model_3f, "lambda")
facCors_RNM <- getmatrix(RNM_model_3f, "sigma_zeta")

pdf(file="/path/RNM.pdf", width=8, height=3)
layout(t(1:2))
RNM_fac_graph <- qgraph.loadings(facLoadings_RNM, theme="colorblind", model="reflective", vsize=c(8,13), asize=5, factorCors=facCors_RNM, groups=groups,color=colors)
RNM_net_graph <- qgraph(net_RNM, theme="colorblind", layout="spring", nodeNames=obsvars, groups=groups,legend.cex=0.2, vsize=8, groups=groups, color=colors)
dev.off()
```

#Network topology indices 

```{r}
#RNM centrality 
centrality_RNM <- centrality_auto(net_RNM)
centrality_RNM_df <- centrality_RNM$node.centrality
centrality_RNM_df <- gather(centrality_RNM_df, measure, value, Strength:ExpectedInfluence, factor_key=T)
centrality_RNM_df$node <- c(1:22)
centrality_RNM_df$model <- rep("RNM",22)

#GGM centrality 
centrality_GGM <- centrality_auto(GGM_net)
centrality_GGM_df <- centrality_GGM$node.centrality
centrality_GGM_df <- gather(centrality_GGM_df, measure, value, Strength:ExpectedInfluence, factor_key=T)
centrality_GGM_df$node <- c(1:22)
centrality_GGM_df$model <- rep("GGM",22)

centrality_df <- rbind(centrality_RNM_df, centrality_GGM_df)

layout(t(1:2))
ggplot(centrality_df,  aes(value,node, color=model))+
  geom_point()+
  geom_path()+
  facet_wrap(~measure, ncol=4)+
  labs(title="Centrality comparison")+
  scale_y_continuous(labels=as.character(centrality_df$node), breaks = centrality_df$node)

#Degree distribution 
GGM_deg <- rowSums(GGM_net !=0)
GGM_deg_df <- data.frame("Node"=c(1:22), "Degree"=GGM_deg)

ggplot(GGM_deg_df, aes(Degree))+
  geom_histogram()

RNM_deg <- rowSums(RNM_net_graph != 0)
RNM_deg_df <- data.frame("Node"=c(1:22), "Degree"=RNM_deg)

ggplot(RNM_deg_df, aes(Degree))+
  geom_histogram()

#Clustering coefficient 
clustcoef_auto(RNM_net_graph)
clustcoef_auto(GGM_net)

layout(t(1:2))
clusteringPlot(RNM_net_graph, signed=TRUE, include="WS")
clusteringPlot(GGM_net, signed=TRUE, include="WS")

```

